Automatically proving equivalence by type-safe reflection

### 2016-01-15 paper-JFP-DTP/content/3-generalProblem.tex
paper : use of listings and corrections

In fact, we don't only want to be able to prove equalities, but more generally equivalence. We want to allow the user the possibility to define his own equivalence notion that interests him, assuming he's able to provide the properties that belongs to the algebraic structure that he wants to use. This equivalence relation on $c$ has the following profile $(\sim=) : (x:c) \rightarrow (y:c) \rightarrow Type$\footnote{This Type would be a Prop in system like Coq that make a distinction between the world of computations and the world of logical statements}, and has to be accompanied by the usual properties of reflexivity, symmetry and transitivity.

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
class Set c where
    (~=) : (x:c) -> (y:c) -> Type

    refl : (x:c) -> x~=x
    sym : {x:c} -> {y:c} -> (x~=y) -> (y~=x)
    trans : {x:c} -> {y:c} -> {z:c} -> (x~=y) -> (y~=z) -> (x~=z)

    set_eq : (x:c) -> (y:c) -> Maybe (x=y)
\end{verbatim}
\end{center}
\caption{Set}
\figrule
\end{figure}
If the user wants to prove propositional equalities, he will simply instantiate $(\sim=)$ with the built-in $(=)$ during the definition of the instance of $Set$.
Note that $(\sim=)$ is only weakly decidable in the sense that $set\_eq$ only produces a proof when the two elements are equivalent, but it doesn't produce a proof of dis-equivalence when they are different -- instead, it simply produces the value $Nothing$--. That's quite natural, since we want to generate proof of equivalence, and not to generate counter examples for proving dis-equivalence, which is another problem.

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
class Set c => Magma c where
    Plus : c -> c -> c
\end{verbatim}
\end{center}
\caption{Magma}
\figrule
\end{figure}

This code means that a type $c$ (for $carrier$) is a Magma if it is already a Set (ie, it is equiped with the equivalence relation $\sim=$ and the test $set\_eq$), and if it has a $Plus$ operation.

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
class Magma c => SemiGroup c where
    Plus_assoc : (c1:c) -> (c2:c) -> (c3:c) 
               -> (Plus (Plus c1 c2) c3 = Plus c1 (Plus c2 c3))
\end{verbatim}
\end{center}
\caption{Semi-Group}
\figrule
\end{figure}

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
class SemiGroup c => Monoid c where
    Zero : c    
    Plus_neutral_1 : (c1:c) -> (Plus Zero c1 = c1)    
    Plus_neutral_2 : (c1:c) -> (Plus c1 Zero = c1)
\end{verbatim}
\end{center}
\caption{Monoid}
\figrule
\end{figure}

A Group is a Monoid, with two new operations. The binary operation $Minus$, and the unary operation $Neg$. We must have the property that $Minus$ can always be simplified with the $Neg$ and the $Plus$. That means that $Minus$ is not a "primitive" operation of a Group, since we can always rewrite $(a-b)$ into $(a\ +\ -b)$. We let the possibility to express a $Minus$ just for convenience for the user, as he won't have to rewrite by hand his $Minus$ operations with sums of negations.
The second axiom, which is the most important one for a Group, is the fact that any value $c1$ admits $Neg\ c1$ for inverse, where being an inverse is the following property :

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
-- This is just a conjunctive predicate
hasSymmetric : (c:Type) -> (p:Monoid c) -> c -> c -> Type
hasSymmetric c p a b = (Plus a b = Zero, Plus b a = Zero)

class Monoid c => Group c where
    Minus : c -> c -> c
    Neg : c -> c
    Minus_simpl : (c1:c) -> (c2:c) -> Minus c1 c2 = Plus c1 (Neg c2) 
    Plus_inverse : (c1:c) -> hasSymmetric c _ c1 (Neg c1)
\end{verbatim}
\end{center}
\caption{Group}
\figrule
\end{figure}

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
data ExprMa : Magma c -> (Vect n c) -> c -> Type where
    ConstMa : (p : Magma c) -> (gam:Vect n c) -> (c1:c)  -> ExprMa p gam c1 
    PlusMa : {p : Magma c} -> {gam:Vect n c} -> {c1:c} -> {c2:c} 
         -> ExprMa p gam c1 -> ExprMa p gam c2 
         -> ExprMa p gam (Plus c1 c2) 
    VarMa : (p:Magma c) -> {gam:Vect n c}
         -> (i:Fin n) -> ExprMa p gam (index i gam)
\end{verbatim}
\end{center}
\caption{Reflected terms for a Magma}
\figrule
\end{figure}

@@

%For the moment, the definition of $Variable$ is just the following:
%
%\begin{code}[caption=Reflected variables, captionpos=b, label=lst1:haskell2]  
%data Variable : {c:Type} -> {n:Nat}
%  -> (c_equal : (c1:c)->(c2:c)->Maybe(c1=c2)) 
%  -> (Vect n c) -> c -> Type where
%    RealVariable : (c_equal:(c1:c)->(c2:c)
%                    ->Maybe(c1=c2))
%          -> (gam:Vect n c) -> (i:Fin n) 
%          -> Variable c_equal gam (index i gam) 
%\end{code}	
%
%This type will be extended with some other constructors later on. For the moment, we can only %express a "real variable", referred by its index $i$. \\
%\\
Because there is no additional operations in a SemiGroup or in a Monoid, the datatypes for reflected terms in these two structures will have exactly the same shape as the one for Magma that we've given above.
However, the one for $Group$ will introduce two new constructors for the $Neg$ and $Minus$ operations :

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
data ExprG :  Group c -> (Vect n c) -> c -> Type where
    ConstG : (p : dataTypes.Group c) -> (gam:Vect n c) 
      -> (c1:c) -> ExprG p gam c1
    PlusG : {p : Group c} -> {gam:Vect n c} -> {c1:c} -> {c2:c} 
      -> ExprG p gam c1 -> ExprG p gam c2 
      -> ExprG p gam (Plus c1 c2)
    MinusG : {p : Group c} -> {gam:Vect n c} -> {c1:c} -> {c2:c} 
      -> ExprG p gam c1 -> ExprG p gam c2 
      -> ExprG p gam (Minus c1 c2)
    NegG : {p : Group c} -> {gam:Vect n c} -> {c1:c} 
      -> ExprG p gam c1 -> ExprG p gam (Neg c1)
    VarG : (p : Group c) -> {gam:Vect n c} 
      -> (i:Fin n) -> ExprG p gam c1
\end{verbatim}
\end{center}
\caption{Reflected terms in a Group}
\figrule
\end{figure}

@@

	\subsection {A bit of notation}
The equality we are trying to prove is $x=y$, where $x$ and $y$ are elements of the type $c$, which  simulates a set with some properties (which make $c$ being a Semi-Group, or a Monoid, or a Group...). The fact that $c$ fulfils the specification of an algebraic structure is expressed as an instance of the corresponding type class, and this instance will be denoted as $p$.
The reflected term for $x$ will be denoted $ex$, and this term will have the type $ExprG\ p\ \Gamma\ x$. The term $ex$ is the encoding of $x$ and its type is precisely indexed over the real value $x$, as the reflected terms embed the concrete values.
We've got similar thing for $y$, which is encoded by $ey$, and its type is indexed over the real value $y$.
Running the normalisation procedure on $ex$ will produce the normal form $ex'$ of type $ExprG\ p\ \Gamma\ x'$ and a proof $px$ of $x=x'$ while running the normalisation procedure on $ey$ will produce the normal form $ey'$ of type $ExprG\ p\ \Gamma\ y'$ and a proof $py$ of $y=y'$.

	\subsection {Deciding equality}

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
	groupReduce : {c:Type} -> {n:Nat} -> (p:Group c) -> {gam:Vect n c} 
	  -> {x:c} -> (ExprG p gam x) -> (x' ** (ExprG p gam x', x=x'))
\end{verbatim}
\end{center}
\caption{Type of the reduction function for terms reflecting elements in a Group}
\figrule
\end{figure}

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
	groupDecideEq : (p:Group c) -> {gam:Vect n c} -> {x : c} -> {y : c} 
	   -> (ExprG p gam x) -> (ExprG p gam y) 
	   -> Maybe (x = y)
\end{verbatim}
\end{center}
\caption{Type of the function for deciding equality between elements in a Group}
\figrule
\end{figure}

The first thing this function has to do is to compute the normal form of the two expressions $ex$ and $ey$ respectively encoding $x$ and $y$.
Then, the only thing remaining to do will be to compare syntactically $ex'$ and $ey'$, and if they are equal, then we will get $x'=y'$, and we will be able to produce the desired proof of $x=y$ by using the two proofs $px$ and $py$.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
groupDecideEq p ex ey =
  let (x' ** (ex', px)) = groupReduce p ex in
  let (y' ** (ey', py)) = groupReduce p ey in
	     buildProofGroup p ex' ey' px py
\end{verbatim}
\end{center}
\caption{Decides if two elements of a group are equal}
\figrule
\end{figure}

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
buildProofGroup : (p:Group c) -> {gam:Vect n c} 
  -> {x : c} -> {y : c} -> {x':c} -> {y':c} 
  -> (ExprG p gam x') -> (ExprG p gam y') 
  -> (x = x') -> (y = y') 
  -> (Maybe (x = y))
buildProofGroup p ex ey px py with (exprG_eq p _ ex ey)
    buildProofGroup p ex ex px py | Just refl = ?MbuildProofGroup
    buildProofGroup p ex ey px py | Nothing = Nothing
\end{verbatim}
\end{center}
\caption{Composes the two proofs if the normal forms are the same}
\figrule
\end{figure}

The proof for the metavariable $MbuildProofGroup$ is exactly the same as the corresponding one (called $MbuildProof$) in the smallest example showed in section 2.3.

\subsection {Automatic reflection}
$reflect : \{n:Nat\} \rightarrow (gamma : Vect\ n\ c) \rightarrow (c1:c) \rightarrow Expr\ gamma\ c1$.
The reflected term is guaranteed to be a faithful representation of the input $c1$. We can write this function because Idris has a $\%syntax$ command which let us do pattern maching on syntax (ie, it let us access the AST of a term), and this $\%syntax$ produces functions which are typed ! (unlike Coq's $Ltac$).

>>>>

In fact, we don't only want to be able to prove equalities, but more generally equivalence. We want to allow the user the possibility to define his own equivalence notion that interests him, assuming he's able to provide the properties that belongs to the algebraic structure that he wants to use. This equivalence relation on $c$ has the following profile $(\simeq) : (x:c) \rightarrow (y:c) \rightarrow Type$\footnote{This Type would be a Prop in system like Coq that make a distinction between the world of computations and the world of logical statements}, and has to be accompanied by the usual properties of reflexivity, symmetry and transitivity.

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{lstlisting}
class Set c where
    ($\simeq$) : (x:c) -> (y:c) -> Type

    refl : (x:c) -> x $\simeq$ x
    sym : {x:c} -> {y:c} -> (x $\simeq$ y) -> (y $\simeq$ x)
    trans : {x:c} -> {y:c} -> {z:c} -> (x $\simeq$ y) -> (y $\simeq$ z) -> (x $\simeq$ z)

    set_eq : (x:c) -> (y:c) -> Maybe (x $\simeq$ y)
\end{lstlisting}
\end{center}
\caption{Set}
\figrule
\end{figure}
If the user wants to prove propositional equalities, he will simply instantiate $(\simeq)$ with the built-in $(=)$ during the definition of the instance of $Set$.
Note that $(\simeq)$ is only weakly decidable in the sense that $set\_eq$ only produces a proof when the two elements are equivalent, but it doesn't produce a proof of dis-equivalence when they are different -- instead, it simply produces the value $Nothing$--. That's quite natural, since we want to generate proof of equivalence, and not to generate counter examples for proving dis-equivalence, which is another problem.

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{lstlisting}
class Set c => Magma c where
    Plus : c -> c -> c
\end{lstlisting}
\end{center}
\caption{Magma}
\figrule
\end{figure}

This code means that a type $c$ (for $carrier$) is a Magma if it is already a Set (ie, it is equiped with the equivalence relation $\simeq$ and the test $set\_eq$), and if it has a $Plus$ operation.

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{lstlisting}
class Magma c => SemiGroup c where
    Plus_assoc : (c1:c) -> (c2:c) -> (c3:c) 
               -> (Plus (Plus c1 c2) c3 $\simeq$ Plus c1 (Plus c2 c3))
\end{lstlisting}
\end{center}
\caption{Semi-Group}
\figrule
\end{figure}

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{lstlisting}
class SemiGroup c => Monoid c where
    Zero : c    
    Plus_neutral_1 : (c1:c) -> (Plus Zero c1 $\simeq$ c1)    
    Plus_neutral_2 : (c1:c) -> (Plus c1 Zero $\simeq$ c1)
\end{lstlisting}
\end{center}
\caption{Monoid}
\figrule
\end{figure}
We can continue the hierarchy with groups, which are a more interesting.
A group is a monoid, with two new operations. The binary operation $Minus$, and the unary operation $Neg$. We must have the property that $Minus$ can always be simplified with the $Neg$ and the $Plus$. That means that $Minus$ is not a "primitive" operation of a Group, since we can always rewrite $(a-b)$ into $(a\ +\ -b)$. We let the possibility to express a $Minus$ just for convenience for the user, as he won't have to rewrite by hand his $Minus$ operations with sums of negations.
The second axiom, which is the most important one for a Group, is the fact that any value $c1$ admits $Neg\ c1$ for inverse, where being an inverse is the following property :

\begin{figure}[H]
\figrule
\begin{center}
\begin{lstlisting}
-- This is just a conjunctive predicate
hasSymmetric : (c:Type) -> (p:Monoid c) -> c -> c -> Type
hasSymmetric c p a b $\simeq$ (Plus a b $\simeq$ Zero, Plus b a $\simeq$ Zero)

class Monoid c => Group c where
    Minus : c -> c -> c
    Neg : c -> c
    Minus_simpl : (c1:c) -> (c2:c) -> Minus c1 c2 $\simeq$ Plus c1 (Neg c2) 
    Plus_inverse : (c1:c) -> hasSymmetric c _ c1 (Neg c1)
\end{lstlisting}
\end{center}
\caption{Group}
\figrule
\end{figure}

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{lstlisting}
data ExprMa : Magma c -> (Vect n c) -> c -> Type where
    ConstMa : (p : Magma c) -> ($\Gamma$:Vect n c) -> (c1:c)  -> ExprMa p $\Gamma$ c1 
    PlusMa : {p : Magma c} -> {$\Gamma$:Vect n c} -> {c1:c} -> {c2:c} 
         -> ExprMa p $\Gamma$ c1 -> ExprMa p $\Gamma$ c2 
         -> ExprMa p $\Gamma$ (Plus c1 c2) 
    VarMa : (p:Magma c) -> {$\Gamma$:Vect n c}
         -> (i:Fin n) -> ExprMa p $\Gamma$ (index i $\Gamma$)
\end{lstlisting}
\end{center}
\caption{Reflected terms for a Magma}
\figrule
\end{figure}

@@

%For the moment, the definition of $Variable$ is just the following:
%
%\begin{code}[caption=Reflected variables, captionpos=b, label=lst1:haskell2]  
%data Variable : {c:Type} -> {n:Nat}
%  -> (c_equal : (c1:c)->(c2:c)->Maybe(c1=c2)) 
%  -> (Vect n c) -> c -> Type where
%    RealVariable : (c_equal:(c1:c)->(c2:c)
%                    ->Maybe(c1=c2))
%          -> ($\Gamma$:Vect n c) -> (i:Fin n) 
%          -> Variable c_equal $\Gamma$ (index i $\Gamma$) 
%\end{code}	
%
%This type will be extended with some other constructors later on. For the moment, we can only %express a "real variable", referred by its index $i$. \\
%\\
Because there is no additional operations in a SemiGroup or in a Monoid, the datatypes for reflected terms in these two structures will have exactly the same shape as the one for Magma that we've given above.
However, the one for $Group$ will introduce two new constructors for the $Neg$ and $Minus$ operations :

\begin{figure}[H]
\figrule
\begin{center}
\begin{lstlisting}
data ExprG :  Group c -> (Vect n c) -> c -> Type where
    ConstG : (p : dataTypes.Group c) -> ($\Gamma$:Vect n c) 
         -> (c1:c) -> ExprG p $\Gamma$ c1
    PlusG : {p : Group c} -> {$\Gamma$:Vect n c} -> {c1:c} -> {c2:c} 
         -> ExprG p $\Gamma$ c1 -> ExprG p $\Gamma$ c2 
         -> ExprG p $\Gamma$ (Plus c1 c2)
    MinusG : {p : Group c} -> {$\Gamma$:Vect n c} -> {c1:c} -> {c2:c} 
         -> ExprG p $\Gamma$ c1 -> ExprG p $\Gamma$ c2 
         -> ExprG p $\Gamma$ (Minus c1 c2)
    NegG : {p : Group c} -> {$\Gamma$:Vect n c} -> {c1:c} 
         -> ExprG p $\Gamma$ c1 -> ExprG p $\Gamma$ (Neg c1)
    VarG : (p : Group c) -> {$\Gamma$:Vect n c} 
         -> (i:Fin n) -> ExprG p $\Gamma$ c1
\end{lstlisting}
\end{center}
\caption{Reflected terms in a Group}
\figrule
\end{figure}

@@

	\subsection {A bit of notation}
The equality we are trying to prove is $x \simeq y$, where $x$ and $y$ are elements of the type $c$, which  simulates a set with some properties (which make $c$ being a Semi-Group, or a Monoid, or a Group...). The fact that $c$ fulfils the specification of an algebraic structure is expressed as an instance of the corresponding type class, and this instance will be denoted as $p$.
The reflected term for $x$ will be denoted $ex$, and this term will have the type $ExprG\ p\ \Gamma\ x$. The term $ex$ is the encoding of $x$ and its type is precisely indexed over the real value $x$, as the reflected terms embed the concrete values.
We've got similar thing for $y$, which is encoded by $ey$, and its type is indexed over the real value $y$.
Running the normalisation procedure on $ex$ will produce the normal form $ex'$ of type $ExprG\ p\ \Gamma\ x'$ and a proof $px$ of $x \simeq x'$ while running the normalisation procedure on $ey$ will produce the normal form $ey'$ of type $ExprG\ p\ \Gamma\ y'$ and a proof $py$ of $y \simeq y'$.

	\subsection {Deciding equivalence}

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{lstlisting}
groupReduce : {c:Type} -> {n:Nat} -> (p:Group c) -> {$\Gamma$:Vect n c} 
	   -> {x:c} -> (ExprG p $\Gamma$ x) -> (x' ** (ExprG p $\Gamma$ x', x $\simeq$ x'))
\end{lstlisting}
\end{center}
\caption{Type of the reduction function for terms reflecting elements in a Group}
\figrule
\end{figure}

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{lstlisting}
groupDecideEq : (p:Group c) -> {$\Gamma$:Vect n c} -> {x : c} -> {y : c} 
	     -> (ExprG p $\Gamma$ x) -> (ExprG p $\Gamma$ y) 
	     -> Maybe (x $\simeq$ y)
\end{lstlisting}
\end{center}
\caption{Type of the function for deciding equality between elements in a Group}
\figrule
\end{figure}

The first thing this function has to do is to compute the normal form of the two expressions $ex$ and $ey$ respectively encoding $x$ and $y$.
Then, the only thing remaining to do will be to compare syntactically $ex'$ and $ey'$, and if they are equal, then we will get $x'=y'$, and we will be able to produce the desired proof of $x \simeq y$ by using the two proofs $px : x \simeq x'$ and $py : y \simeq y'$.

\begin{figure}[H]
\figrule
\begin{center}
\begin{lstlisting}
groupDecideEq p ex ey =
  let (x' ** (ex', px)) = groupReduce p ex in
  let (y' ** (ey', py)) = groupReduce p ey in
	     buildProofGroup p ex' ey' px py
\end{lstlisting}
\end{center}
\caption{Decides if two elements of a group are equal}
\figrule
\end{figure}

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{lstlisting}
buildProofGroup : (p:Group c) -> {$\Gamma$:Vect n c} 
  -> {x : c} -> {y : c} -> {x':c} -> {y':c} 
  -> (ExprG p $\Gamma$ x') -> (ExprG p $\Gamma$ y') 
  -> (x $\simeq$ x') -> (y $\simeq$ y') 
  -> (Maybe (x $\simeq$ y))
buildProofGroup p ex ey px py with (exprG_eq p _ ex ey)
    buildProofGroup p ex ex px py | Just refl = ?MbuildProofGroup
    buildProofGroup p ex ey px py | Nothing = Nothing
\end{lstlisting}
\end{center}
\caption{Composes the two proofs if the normal forms are the same}
\figrule
\end{figure}
We've used here the function $exprG\_eq$ to decide the syntactical equality between the reflected terms. This function plays the same role as $eqExpr$, presented on section 2.3 for the smaller example. Also, the proof for the metavariable $MbuildProofGroup$ is exactly the same as the corresponding one (called $MbuildProof$) for the smaller example.

\subsection {Automatic reflection}

\begin{figure}[H]
\figrule
\begin{center}
\begin{lstlisting}
reflect : 
\end{lstlisting}
\end{center}
\caption{Type of the reflecting function for ...}
\figrule
\end{figure}

The reflected term is guaranteed to be a faithful representation of the input $c1$. We can write this function because Idris has a $\%syntax$ command which let us do pattern maching on syntax (ie, it let us access the AST of a term), and this $\%syntax$ produces functions which are typed ! (unlike Coq's $Ltac$).

### 2016-01-15 paper-JFP-DTP/content/4-detailsNormalisation.tex
paper : use of listings and corrections

We now describe briefly how the normalisation functions work for some of the different algebraic structures, and we also explain how we've been able to reuse each prover for the more specialised underneath structures. By doing so, we show that the construction of the desired proof is always done bit by bit, because every rewriting of the reflected term is immediately justified by a little accompanying proof which shows the equivalence of the previous and new index.
\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
littleRewriting : (p:algebraic structure on c) -> (gam:Vect n c) -> {c1:c} 
        -> (Expr p gam c1) -> (c2 ** (Expr p gam c2, c1=c2))
simpleTreatment p gam e = (new expression, justification)
\end{verbatim}
\end{center}
\caption{General pattern of a function doing a part of the normalisation}
\figrule
\end{figure}

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
assoc : (p:SemiGroup c) -> (gam:Vect n c) -> {c1:c} 
        -> (ExprSG p gam c1) -> (c2 ** (ExprSG p gam c2, c1=c2))
assoc p gam (PlusSG 
             (PlusSG ex (ConstSG _ _ const1)) 
             (PlusSG (ConstSG _ _ const2) ey)) =
    let (r_ihx ** (e_ihx, p_ihx)) = (assoc p gam ex) in
    let (r_ihy ** (e_ihy, p_ihy)) = (assoc p gam ey) in
    let (r_3 ** (e_3, p_3)) = magmaReduce (semiGroup_to_magma 
                             (PlusSG (ConstSG _ _ const1) (ConstSG _ _ const2))) in
    let e_3' = magma_to_semiGroup p e_3 in
        (_ ** ((PlusSG (PlusSG e_ihx e_3') e_ihy), ?Massoc1))
[...]
\end{verbatim}
\end{center}
\caption{Computing with associativity in a Semi-Group, first pattern}
\figrule
\end{figure}
The treatment is applied recursively on the sub-expressions $ex$ and $ey$, which produces the new expressions $e\_ihx$ and $e\_ihy$ together with the new concrete values they represent ($r\_ihx$ and $r\_ihy$) and the proofs ($p\_ihx$ and $p\_ihy$) justifying these transformation on the concrete values. Note that the simplification $[const1+const2]$ is done by calling the magma prover, which gives a new expression $e3$, even thought it could perfectly be done directly here. We wanted to show how a prover can be reused for building another prover on an easy example. The functions of conversion between the different levels : $semiGroup\_to\_magma$ and $magma\_to\_semiGroup$ are helping for this task.
\\
A meta-variable Massoc1 has been introduced. This metavariable corresponds to a proof obligation for the following property :
$Massoc1 : Plus\ (Plus\ x\ const1)\ (Plus\ const2\ y)\ ~=\ Plus\ (Plus\ r\_ihx\ r_3)\ r\_ihy$ in a context where there is, amongst other things, $p\_ihx : x\ ~=\ r\_ihx$, $p\_ihy : y\ ~=\ r\_ihy$ and $p_3 : Plus\ const1\ const2 ~= r\_3$.
This lemma says that the treatment done by this function $assoc$ on this case was sound because the previous and the new concrete values are equal. It can easily be proven with a few rewritings and the use of the associativity, available from the semi-group structure. \\
\\
After this simplification of constants computable thanks to associativity, the normalization for SemiGroup also has to rearrange the brackets on a systematic way. The fully right or fully left associative forms can both be chosen for this purpose, and the implementation is very similar to the function $expr_l$ previously described for the smaller example on section 2.3.

@@

%\begin{figure}[H]
%\figrule
%\begin{center}
%\begin{verbatim}
%elimZero : (p:Monoid c)->(gam:Vect n c)->{c1:c} 
%            -> (ExprMo p gam c1) 
%            -> (c2 ** (ExprMo p gam c2, c1=c2))
%elimZero c p gam (PlusMo (ConstMo _ _ const1) (VarMo _ v)) 
%        with (eq_on_set p Zero const1)
%   elimZero c p gam (PlusMo (ConstMo _ _ Zero) (VarMo _ v)) 
%        | (Just refl) = (_ ** (VarMo _ v, ?MelimZero1))
%   elimZero c p gam (PlusMo (ConstMo _ _ const1) (VarMo _ v)) 
%        | _ = (_ ** (PlusMo (ConstMo _ _ const1) (VarMo _ v), refl)) 
%elimZero c p gam 
% (PlusMo (VarMo _ v) (ConstMo _ _ const2)) 
%      with (eq_as_elem_of_set p Zero const2) 
%   elimZero c p gam (PlusMo (VarMo _ v) 
%            (ConstMo _ _ Zero)) | (Just refl) 
%          = (_ ** (VarMo _ _ v, ?MelimZero2))
%   elimZero c p gam (PlusMo (VarMo _ v) 
%            (ConstMo _ _ const2)) | _ 
%          = (_ ** (PlusMo (VarMo _ v) 
%                 (ConstMo _ _ const2), refl))
%[...]
%\end{verbatim}
%\end{center}
%\caption{Eliminating zeros in a Monoid, first pattern}
%\figrule
%\end{figure}

>>>>

We now describe briefly how the normalisation functions work for some of the different algebraic structures, and we also explain how we've been able to reuse each prover for the more specialised underneath structures. By doing so, we show that the construction of the desired proof is always done bit by bit, because every rewriting of the reflected term is immediately justified by a little accompanying proof which shows the equivalence of the previous and new index.
\begin{figure}[H]
\figrule
\begin{center}
\begin{lstlisting}
littleRewriting : (p:algebraic structure on c) -> (gam:Vect n c) -> {c1:c} 
        -> (Expr p gam c1) -> (c2 ** (Expr p gam c2, c1 $\simeq$ c2))
simpleTreatment p gam e = (new expression, justification)
\end{lstlisting}
\end{center}
\caption{General pattern of a function doing a part of the normalisation}
\figrule
\end{figure}

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{lstlisting}
assoc : (p:SemiGroup c) -> (gam:Vect n c) -> {c1:c} 
        -> (ExprSG p gam c1) -> (c2 ** (ExprSG p gam c2, c1 $\simeq$ c2))
assoc p gam (PlusSG 
             (PlusSG ex (ConstSG _ _ const1)) 
             (PlusSG (ConstSG _ _ const2) ey)) =
    let (r_ihx ** (e_ihx, p_ihx)) = (assoc p gam ex) in
    let (r_ihy ** (e_ihy, p_ihy)) = (assoc p gam ey) in
    let (r_3 ** (e_3, p_3)) = magmaReduce (semiGroup_to_magma 
                             (PlusSG (ConstSG _ _ const1) (ConstSG _ _ const2))) in
    let e_3' = magma_to_semiGroup p e_3 in
        (_ ** ((PlusSG (PlusSG e_ihx e_3') e_ihy), ?Massoc1))
[...]
\end{lstlisting}
\end{center}
\caption{Computing with associativity in a Semi-Group, first pattern}
\figrule
\end{figure}
The treatment is applied recursively on the sub-expressions $ex$ and $ey$, which produces the new expressions $e\_ihx$ and $e\_ihy$ together with the new concrete values they represent ($r\_ihx$ and $r\_ihy$) and the proofs ($p\_ihx$ and $p\_ihy$) justifying these transformation on the concrete values. Note that the simplification $[const1+const2]$ is done by calling the magma prover, which gives a new expression $e3$, even thought it could perfectly be done directly here, but we wanted to show how a prover can be reused for building another prover on an easy example. The functions of conversion between the different levels : $semiGroup\_to\_magma$ and $magma\_to\_semiGroup$ are helping for this task.
\\
A meta-variable Massoc1 has been introduced. This metavariable corresponds to a proof obligation for the following property :
$Massoc1 : Plus\ (Plus\ x\ const1)\ (Plus\ const2\ y)\ \simeq\ Plus\ (Plus\ r\_ihx\ r_3)\ r\_ihy$ in a context where there is, amongst other things, $p\_ihx : x\ \simeq\ r\_ihx$, $p\_ihy : y\ \simeq\ r\_ihy$ and $p_3 : Plus\ const1\ const2 \simeq r\_3$.
This lemma says that the treatment done by this function $assoc$ on this pattern was sound because the previous and the new concrete values are equal. It can easily be proven with a few rewritings and the use of the associativity, available from the semi-group structure. \\
\\
After this simplification of constants computable thanks to associativity, the normalization for SemiGroup also has to rearrange the brackets on a systematic way. The fully right or fully left associative forms can both be chosen for this purpose, and the implementation is very similar to the function $expr_l$ previously described for the smaller example on section 2.3.

@@

%\begin{figure}[H]
%\figrule
%\begin{center}
%\begin{lstlisting}
%elimZero : (p:Monoid c)->(gam:Vect n c)->{c1:c} 
%            -> (ExprMo p gam c1) 
%            -> (c2 ** (ExprMo p gam c2, c1 $\simeq$ c2))
%elimZero c p gam (PlusMo (ConstMo _ _ const1) (VarMo _ v)) 
%        with (eq_on_set p Zero const1)
%   elimZero c p gam (PlusMo (ConstMo _ _ Zero) (VarMo _ v)) 
%        | (Just refl) = (_ ** (VarMo _ v, ?MelimZero1))
%   elimZero c p gam (PlusMo (ConstMo _ _ const1) (VarMo _ v)) 
%        | _ = (_ ** (PlusMo (ConstMo _ _ const1) (VarMo _ v), refl)) 
%elimZero c p gam 
% (PlusMo (VarMo _ v) (ConstMo _ _ const2)) 
%      with (eq_as_elem_of_set p Zero const2) 
%   elimZero c p gam (PlusMo (VarMo _ v) 
%            (ConstMo _ _ Zero)) | (Just refl) 
%          = (_ ** (VarMo _ _ v, ?MelimZero2))
%   elimZero c p gam (PlusMo (VarMo _ v) 
%            (ConstMo _ _ const2)) | _ 
%          = (_ ** (PlusMo (VarMo _ v) 
%                 (ConstMo _ _ const2), refl))
%[...]
%\end{lstlisting}
%\end{center}
%\caption{Eliminating zeros in a Monoid, first pattern}
%\figrule
%\end{figure}
